\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{times}
\usepackage[title, titletoc]{appendix}
\usepackage{graphicx}
\usepackage{amsmath, amssymb}
\usepackage{mathtools} % for \xmapsto
\usepackage{bm}
\usepackage[most]{tcolorbox}
\usepackage{tabularx}
\usepackage{pifont} % optional for symbols
% --- Flowcharts (TikZ) ---
\usepackage{tikz}
\usetikzlibrary{arrows.meta,positioning,shapes.geometric,fit,calc}
\usepackage{xcolor}
% Algorithms (compact)
\usepackage{algorithm}
\usepackage{algpseudocode}

% Optional: slightly tighter algorithm typography
\algrenewcommand\algorithmicrequire{\textbf{Input:}}
\algrenewcommand\algorithmicensure{\textbf{Output:}}
\algrenewcommand\algorithmiccomment[1]{\hfill{\footnotesize$\triangleright$~#1}}

\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{booktabs,longtable,array,makecell}
\setlength{\LTleft}{0pt}   % no left indent for longtable
\setlength{\LTright}{0pt}  % no right indent for longtable
\newcolumntype{C}[1]{>{\centering\arraybackslash}p{#1}}
\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}


% --- Theorems, lemmas, claims, proofs ---
\usepackage{amsthm}

\theoremstyle{plain} % bold headings, italic body
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{claim}[theorem]{Claim}

\theoremstyle{definition} % bold heading, upright body
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{example}[theorem]{Example}

\theoremstyle{remark} % italic heading, upright body
\newtheorem{remark}[theorem]{Remark}
\newtheorem{note}[theorem]{Note}

% Optional: unnumbered versions
\newtheorem*{theorem*}{Theorem}
\newtheorem*{lemma*}{Lemma}
\newtheorem*{claim*}{Claim}

% Proof environment is provided by amsthm:
% \begin{proof} ... \end{proof}
% Customize QED symbol if you like:
\renewcommand{\qedsymbol}{$\blacksquare$}

\newcommand{\code}[1]{\texttt{#1}}

\definecolor{prefillcolor}{RGB}{173,216,230}  % Light blue
\definecolor{decodecolor}{RGB}{144,238,144}   % Light green
\definecolor{gridcolor}{RGB}{200,200,200}     % Light gray
\definecolor{highlightcolor}{RGB}{255,182,193} % Light pink for integration region


% (Optional) If you want equations numbered by section (since you already use amsmath):
\numberwithin{equation}{section}



\title{BLIS: A Blackbox Inference Simulator for Capacity Planning,\\
Configuration Search, and Algorithm Evolution}


\author{AI Platform Optimization Team}
\date{\today}

\begin{document}

\maketitle

% ============================================================
% ABSTRACT
% ============================================================
\begin{abstract}

% [Para 1: Problem and gap]
% Operators of large-scale LLM inference platforms must answer three
% questions daily: How many GPUs do I need? What serving configuration
% minimizes tail latency? And can a new routing or scheduling algorithm
% improve my fleet? Today, answering any of these requires expensive
% real-hardware experiments or simulators that sacrifice either fidelity
% or generality.

% [Para 2: BLIS approach]
% We present BLIS, a blackbox discrete-event simulator for LLM inference
% serving. BLIS factorizes inference performance into learned execution
% physics, pluggable control-plane policies, and composable workload
% models. It requires no access to engine internals: 10 coefficients
% fitted from production traces suffice to predict step-level GPU
% execution with 7\% MAPE across four model architectures. BLIS models
% the full serving stack---admission, routing, scheduling, KV cache
% tiering, and batch formation---and runs 1000x faster than real
% deployments on commodity CPUs.

% [Para 3: Contributions and results]
% We evaluate BLIS on three operator problems. (1) Accuracy: BLIS
% predicts p50/p95/p99 latency within X\% of production measurements,
% outperforming Vidur, InferSym, BentoML Optimizer, and NVIDIA AI
% Configurator by Y--Z\%. (2) Capacity planning and configuration
% search: BLIS correctly identifies GPU counts and serving configurations
% that satisfy SLO targets, avoiding the over-provisioning and
% under-provisioning errors made by competing tools. (3) Algorithm
% evolution: using BLIS as a fitness function, we evolve routing and
% scheduling policies via AI-native systems (ADRS) methodology and
% transfer winning strategies to production, achieving X\% throughput
% improvement with no SLO violations. All insights are validated through
% sim-to-real transfer experiments on H100 clusters.

\end{abstract}


% ============================================================
% SECTION 1: INTRODUCTION
% ============================================================
\section{Introduction}
\label{sec:intro}

% --- The three operator questions ---
% Open with the daily reality of LLM platform operators: capacity
% planning ("how many GPUs?"), configuration search ("what TP/batch/cache
% settings?"), and algorithm selection ("which routing policy?"). Each
% is currently answered by expensive real-hardware experiments.

% --- Why existing tools fall short ---
% Briefly categorize existing approaches: (1) analytical models that
% ignore queueing and batching dynamics, (2) trace-replay simulators
% (Vidur) that hard-code engine internals and break on new configurations,
% (3) vendor-specific optimizers (NVIDIA AI Configurator, BentoML) that
% treat the engine as a black box but lack serving-stack fidelity.
% State the gap: no existing tool provides the combination of
% blackbox portability, serving-stack fidelity, and speed needed for
% all three operator problems.

% --- BLIS in one paragraph ---
% Introduce BLIS: a CPU-only, deterministic, discrete-event simulator
% that factorizes inference performance into learned execution physics
% (10 coefficients from production traces), pluggable control-plane
% policies (admission, routing, scheduling, KV cache, batch formation),
% and composable workload models. Emphasize the blackbox property: BLIS
% requires no source-code access to the inference engine.

% --- Key insight ---
% State the central insight: the performance-critical mechanisms of an
% inference engine (step time, memory transfer latency) are low-dimensional
% and learnable from sparse production traces, while the control-plane
% policies (routing, scheduling, admission) are deterministic and can be
% faithfully replicated. This separation enables a simulator that is both
% accurate and fast enough to serve as a fitness function for automated
% algorithm search.

% --- Contributions (numbered list) ---
% C1: BLIS simulator with factorized architecture
% C2: Trained-roofline latency model (7% MAPE, 4 architectures, 137K requests)
% C3: Accuracy comparison against Vidur, InferSym, BentoML, NVIDIA
% C4: Capacity planning and configuration search case studies with sim-to-real
% C5: Algorithm evolution via ADRS with sim-to-real transfer
% C6: Open-source release of simulator, training pipeline, and trace dataset

% --- Paper roadmap ---
% Brief forward-reference to sections.


% ============================================================
% SECTION 2: BACKGROUND AND RELATED WORK
% ============================================================
\section{Background and Related Work}
\label{sec:background}

This section establishes the technical context for BLIS.
We first describe the architecture of modern LLM inference serving systems (\S\ref{subsec:bg-serving}), then survey the landscape of performance prediction tools (\S\ref{subsec:bg-simulators}), workload characterization (\S\ref{subsec:bg-workloads}), capacity planning practices (\S\ref{subsec:bg-capplan}), disaggregated serving architectures (\S\ref{subsec:bg-disagg}), and AI-driven algorithm design (\S\ref{subsec:bg-ainative}).
For each area, we identify the specific gap that motivates BLIS.

% -------------------------------------------------------
\subsection{LLM Inference Serving Architecture}
\label{subsec:bg-serving}

Modern LLM inference engines~\cite{kwon2023vllm, yu2022orca} execute requests through a two-phase pipeline.
In the \emph{prefill} phase, the engine processes all input tokens in a single (or chunked) forward pass to populate the key-value (KV) cache.
In the \emph{decode} phase, the engine autoregressively generates output tokens, one per forward pass, reading from and appending to the KV cache.
These two phases have fundamentally different computational profiles: prefill is compute-bound (high arithmetic intensity), while decode is memory-bandwidth-bound (low reuse of loaded weights)~\cite{yuan2024llminference, williams2009roofline}.

The serving engine orchestrates requests through a \emph{busy-loop scheduler} that constructs a batch of active requests at each iteration (called a \emph{step}).
Continuous batching~\cite{yu2022orca} allows new requests to join and completed requests to leave the batch at every step, rather than waiting for an entire batch to finish.
Chunked prefill~\cite{agrawal2023sarathi, agrawal2024sarathiserve} further interleaves prefill and decode within a single step by splitting large prompts into fixed-size chunks, reducing head-of-line blocking and improving GPU utilization.

\paragraph{KV cache management.}
The KV cache is the dominant memory consumer in LLM serving.
PagedAttention~\cite{kwon2023vllm} manages KV blocks in a paged virtual-memory style, enabling dynamic allocation and reducing fragmentation.
Multi-tier KV caching extends this to CPU DRAM and SSD~\cite{sheng2023flexgen, qin2024mooncake}, offloading evicted blocks and reloading them when needed---at the cost of transfer latency.
Prefix caching~\cite{zheng2023sglang} reuses KV blocks across requests that share common prompt prefixes, amortizing prefill cost for workloads with high prefix locality (e.g., system prompts, few-shot examples).

\paragraph{Key metrics.}
We adopt the standard metrics used by operators:
\emph{Time to first token} (TTFT) measures the latency from request arrival to the first generated token.
\emph{Time per output token} (TPOT) measures the inter-token latency during decode.
\emph{End-to-end latency} (E2E) is the total request duration.
Operators define \emph{service-level objectives} (SLOs) on tail percentiles (p95, p99) of these metrics.

% -------------------------------------------------------
\subsection{Inference Performance Prediction}
\label{subsec:bg-simulators}

Predicting the performance of an LLM serving deployment \emph{before} running it on real GPUs is a central challenge for capacity planning.
Existing approaches span a spectrum from analytical models to full-fidelity simulation.

\paragraph{Analytical and roofline models.}
The roofline model~\cite{williams2009roofline} bounds achievable performance by the minimum of peak compute throughput and memory bandwidth, parameterized by a kernel's arithmetic intensity.
Recent work applies roofline analysis to transformer inference~\cite{yuan2024llminference}, yielding useful first-order estimates of per-step GPU time.
However, pure roofline models do not capture queueing dynamics, batching interference, KV cache pressure, or multi-instance routing---all of which dominate end-to-end latency in production.

\paragraph{Vidur.}
Vidur~\cite{agrawal2024vidur} is the most closely related prior work.
It is a large-scale simulation framework that models LLM inference by profiling individual GPU operators and composing their execution times into step-level predictions.
Vidur achieves less than 9\% error in single-instance latency estimation and includes Vidur-Search, a configuration optimizer that evaluates deployment options on CPU.
However, Vidur's profiling approach requires per-operator measurements on the target GPU, tightly coupling the simulator to a specific engine version and hardware platform.
When the engine's scheduler changes (e.g., new batching policy) or when extrapolating to unseen configurations (different TP degree, different KV cache sizes), the operator profiles may no longer be valid.
BLIS takes a fundamentally different approach: rather than profiling engine operators, BLIS learns a compact step-time model from end-to-end traces, decoupling prediction accuracy from engine internals.

\paragraph{InferSim.}
InferSim~\cite{alibaba2025infersim}, also from Alibaba, is a lightweight pure-Python analytical simulator that predicts TTFT, TPOT, and throughput from FLOPs calculations and empirical kernel benchmarks (MFU measurements from FlashAttention, FlashMLA, DeepGEMM).
InferSim targets \emph{model-system co-design}: given model hyperparameters and hardware specs, it identifies whether a configuration is compute-bound or memory-bandwidth-bound.
Its predictions are accurate for single-instance steady-state throughput (within 10--15\% on DeepSeek-V3 and Qwen3-30B), but it does not model queueing, batching dynamics, multi-instance routing, or KV cache pressure---the serving-stack phenomena that dominate end-to-end latency under production load.
InferSim and BLIS are complementary: InferSim answers ``how fast is one step on this hardware?'' while BLIS answers ``what happens to p95 latency when 1{,}000 requests/second hit a 4-instance cluster with this routing policy?''

\paragraph{ML system simulators for training.}
Simulation-based performance estimation is well-established for ML \emph{training} workloads.
ASTRA-sim~\cite{won2023astrasim2} models hierarchical networks and disaggregated systems for distributed training.
Phantora~\cite{qin2025phantora} (NSDI~'26) takes a hybrid approach, executing unmodified training frameworks in containerized environments while intercepting and simulating GPU operations using a single physical GPU.
These training simulators validate the methodology of simulation-based estimation for ML systems, but they do not model inference-specific phenomena: request-level queueing, continuous batching, KV cache dynamics, or multi-instance routing.
BLIS brings analogous simulation fidelity to the inference serving domain.

\paragraph{Vendor and platform tools.}
NVIDIA provides TensorRT-LLM~\cite{nvidia2024tensorrtllm} for optimizing inference on NVIDIA GPUs, and NVIDIA NIM~\cite{nvidia2024nim} includes configuration guidance for deployment.
BentoML~\cite{bentoml2024} provides a deployment framework for LLM serving with built-in batching and parallelism support.
These tools optimize the \emph{runtime} for a specific deployment but do not offer predictive simulation: they cannot answer ``what if'' questions about configurations or workloads not yet deployed.
In contrast, BLIS simulates the full serving stack---including routing, admission control, and KV cache dynamics---enabling counterfactual analysis without deploying to real GPUs.

\paragraph{Cost and efficiency analysis.}
Several works analyze LLM inference cost and efficiency.
Narayanan et al.~\cite{narayanan2023cheaply} propose metrics for comparing inference efficiency across API providers.
Surveys by Zhou et al.~\cite{zhou2024efficientinference} and Li et al.~\cite{li2024llmservingsurvey} catalog optimization techniques across model compression, algorithmic improvements, and system-level enhancements.
These provide useful taxonomies but do not offer simulation tools for predicting performance under specific configurations and workloads.

\paragraph{Summary of the gap.}
Table~\ref{tab:simulator-comparison} summarizes the capabilities of existing prediction approaches.
No existing tool simultaneously provides: (1) blackbox operation (no engine source access required), (2) full serving-stack simulation (routing, admission, scheduling, KV cache), (3) transferability across configurations without re-profiling, and (4) sufficient speed to serve as a fitness function for automated search.
BLIS fills this gap.

\begin{table}[t]
\centering
\small
\caption{Comparison of inference performance prediction approaches. \textbf{Blackbox}: does not require engine source code or per-operator profiling. \textbf{Serving stack}: models multi-instance routing, admission control, and scheduling. \textbf{KV dynamics}: models KV cache pressure, tiering, and prefix reuse. \textbf{Transferable}: accurate under configuration changes without re-profiling. \textbf{Search speed}: fast enough for automated configuration search ($\ge$100 evaluations/hour on CPU).}
\label{tab:simulator-comparison}
\begin{tabular}{lccccc}
\toprule
\textbf{Approach} & \textbf{Blackbox} & \textbf{Serving stack} & \textbf{KV dynamics} & \textbf{Transferable} & \textbf{Search speed} \\
\midrule
Roofline~\cite{williams2009roofline, yuan2024llminference} & \ding{51} & \ding{55} & \ding{55} & \ding{51} & \ding{51} \\
Vidur~\cite{agrawal2024vidur} & \ding{55} & Partial & Partial & \ding{55} & \ding{51} \\
InferSim~\cite{alibaba2025infersim} & \ding{51} & \ding{55} & \ding{55} & \ding{51} & \ding{51} \\
BentoML LLMOptimizer~\cite{bentoml2024} & \ding{51} & \ding{55} & \ding{55} & \ding{55} & \ding{51} \\
\textbf{BLIS-analytical} & \ding{51} & \ding{51} & \ding{51} & \ding{51} & \ding{51} \\
\textbf{BLIS-trained} & \ding{55} & \ding{51} & \ding{51} & \ding{51} & \ding{51} \\
\bottomrule
\end{tabular}
\end{table}

% -------------------------------------------------------
\subsection{Workload Characterization and Generation}
\label{subsec:bg-workloads}

Realistic workload modeling is critical for simulation fidelity.
Production LLM serving workloads exhibit complex structure: multi-turn conversations with varying inter-turn gaps, heterogeneous request sizes (from short chat queries to long-context RAG), bursty arrival patterns driven by user behavior, and prefix locality from shared system prompts~\cite{xiang2025servegen}.

ServeGen~\cite{xiang2025servegen} provides the most comprehensive characterization to date, analyzing production traces from Alibaba's cloud inference service across language, multimodal, and reasoning models.
ServeGen composes workloads on a per-client basis, capturing the correlation between arrival patterns and request characteristics---an approach that avoids the 50\% under-provisioning errors of naive workload generation.
BLIS's workload model (\S\ref{subsec:design-workloads}) draws directly on ServeGen's per-client compositional design, supporting multi-client specifications with configurable arrival processes, token distributions, SLO classes, and cohort dynamics.

Benchmarking tools such as inference-perf~\cite{inferenceperf2024} generate controlled load for measuring real-system performance.
We use inference-perf both for BLIS's training data collection and for sim-to-real validation experiments.

% -------------------------------------------------------
\subsection{Capacity Planning and Configuration Search}
\label{subsec:bg-capplan}

Deploying an LLM serving system requires choosing values for a large number of interdependent parameters: model variant, GPU type, tensor parallelism (TP) degree, maximum batch size, chunked-prefill budget, KV cache tier sizes, routing policy, and admission control thresholds.
The configuration space grows combinatorially, and interactions between parameters are non-obvious: for example, increasing TP degree reduces per-step latency but increases communication overhead, and the net effect depends on the workload's prefill-to-decode ratio.

Today, operators navigate this space through manual benchmarking on real hardware, often guided by vendor recommendations or rules of thumb.
Vidur-Search~\cite{agrawal2024vidur} automates part of this process by simulating candidate configurations on CPU, finding the optimal LLaMA-2-70B configuration in one hour versus the estimated 42{,}000 GPU-hours of exhaustive search.
However, Vidur-Search inherits Vidur's profiling-based approach, limiting its applicability to configurations for which operator profiles exist.

BLIS extends this capability in two ways.
First, because BLIS's learned execution model generalizes across configurations (\S\ref{subsubsec:trained-roofline}), it can evaluate configurations outside the training distribution without re-profiling.
Second, BLIS simulates the full serving stack---including multi-instance routing and admission control---enabling optimization over serving-layer parameters that Vidur-Search does not model.

% -------------------------------------------------------
\subsection{Disaggregated and Distributed Serving}
\label{subsec:bg-disagg}

A growing body of work separates the prefill and decode phases onto distinct hardware to exploit their differing computational profiles.
DistServe~\cite{zhong2024distserve} disaggregates prefill and decode onto separate GPUs, achieving 7.4$\times$ higher throughput by independently optimizing resource allocation for each phase.
Splitwise~\cite{patel2024splitwise} splits inference phases across machines optimized for each workload, transferring KV cache state over GPU cluster interconnects to achieve 1.4$\times$ throughput at 20\% lower cost.
Mooncake~\cite{qin2024mooncake} takes disaggregation further with a KVCache-centric architecture that repurposes idle CPU, DRAM, and SSD capacity as a distributed cache layer, enabling 75\% more requests under production conditions.
MemServe~\cite{hu2024memserve} integrates context caching with disaggregated inference through an elastic memory pool.

These systems introduce architectural complexity---cross-node KV transfer, cache-aware routing, phase-specific scheduling---that simulators must model to produce accurate predictions.
BLIS provides the necessary abstractions: tiered KV cache with configurable transfer bandwidth, per-instance schedulers, and cluster-level routing with composable scoring functions.
The llm-d project~\cite{llmd2024} embodies these ideas in a Kubernetes-native serving stack with disaggregated prefill/decode, tiered KV caching, and intelligent routing; BLIS's default routing profile (prefix-affinity:3, queue-depth:2, kv-utilization:2) is designed for parity with llm-d's endpoint picker.

\paragraph{Other serving optimizations.}
Orca~\cite{yu2022orca} introduced iteration-level scheduling for continuous batching.
FastServe~\cite{wu2023fastserve} uses preemptive scheduling with a multi-level feedback queue.
AlpaServe~\cite{li2023alpaserve} exploits statistical multiplexing with model parallelism to handle bursty workloads.
SGLang~\cite{zheng2023sglang} accelerates structured generation with RadixAttention for KV cache reuse.
S-LoRA~\cite{sheng2023slora} and Punica~\cite{chen2023punica} enable multi-tenant serving of many LoRA adapters.
InferCept~\cite{abhyankar2024infercept} optimizes serving for augmented LLMs that interact with external tools.
LoongServe~\cite{wu2024loongserve} introduces elastic sequence parallelism for long-context models.
Each of these systems introduces scheduling, caching, or parallelism innovations that affect performance in ways existing simulators do not capture.
BLIS's pluggable policy architecture (\S\ref{subsec:design-policies}) provides the extensibility to incorporate these innovations as new policy modules.

% -------------------------------------------------------
\subsection{AI-Native Systems and Algorithm Evolution}
\label{subsec:bg-ainative}

A recent line of work uses large language models to \emph{design and evolve} algorithms, rather than merely execute them.
FunSearch~\cite{romeraparedes2024funsearch} pairs an LLM with an automated evaluator to iteratively evolve programs that solve open mathematical problems, discovering new solutions for the cap set problem and more efficient bin-packing algorithms.
EvoPrompting~\cite{chen2024evoprompting} uses LLMs for code-level neural architecture search.
Lehman et al.~\cite{lehman2023evolution} demonstrate evolution through large models, using LLMs as mutation operators in evolutionary algorithms.

These approaches share a common requirement: a fast, accurate evaluation function that can score thousands of candidate solutions.
In the context of LLM serving, the ``algorithms'' to be evolved are routing policies, scheduling strategies, and admission control parameters---and the evaluation function must simulate their end-to-end impact on latency, throughput, and SLO compliance under realistic workloads.

Real-hardware evaluation is prohibitively expensive for evolutionary search (each configuration evaluation requires minutes of GPU time).
BLIS enables this methodology by providing a simulator that is both faithful enough to predict real-system behavior (\S\ref{sec:eval-accuracy}) and fast enough to evaluate hundreds of candidate policies per hour on commodity CPUs (\S\ref{sec:eval-evolution}).
We demonstrate this capability concretely through the ADRS (Automated Design of Routing Strategies) methodology in Section~\ref{sec:eval-evolution}.


% ============================================================
% SECTION 3: SIMULATING INFERENCE IS HARDER THAN YOU THINK
% ============================================================
\section{Simulating Inference Is Harder Than You Think}
\label{sec:hard}

(To be written: problem-depth section showing with measurements that existing simulators fail in specific, quantifiable ways.)

\subsection{Measurement Setup}
\label{subsec:hard-setup}
\subsection{Step Time Prediction Is Not Enough}
\label{subsec:hard-steptime}
\subsection{Configuration Extrapolation Breaks Existing Tools}
\label{subsec:hard-extrap}
\subsection{The Capacity Planning Error Cascade}
\label{subsec:hard-cascade}

% ============================================================
% SECTION 4: BLIS DESIGN
% ============================================================
\section{BLIS: Blackbox Inference Simulator}
\label{sec:design}

Serving a large language model requires reasoning simultaneously about GPU execution
physics, cluster-wide control-plane decisions, and time-varying workload behavior.
In a running system these concerns are entangled: a routing policy change reshapes
KV-cache hit rates, which alters batch composition, which changes step time.
Yet they are \emph{analytically separable}, and that separation is the design center
of BLIS.

\paragraph{Design principle.}
BLIS factorizes inference simulation into three independent, swappable axes
(Figure~\ref{fig:blis-arch}):
(1)~\emph{learned execution physics}---one of four interchangeable latency-model
backends predicts how long a batch step takes;
(2)~\emph{pluggable control-plane policies}---admission, routing, scheduling, and
memory management replicate the deterministic serving logic faithfully in software,
without GPUs;
and~(3)~\emph{composable workload models}---a declarative specification generates
realistic request streams without a live deployment.
Any axis can be replaced without touching the other two.
This factorization makes BLIS \emph{portable} (no engine source code required),
\emph{accurate} (7\% MAPE in step-time prediction across four architectures,
\S\ref{subsubsec:training}), and \emph{fast} (a 1{,}000-request simulation
completes in under two seconds on a commodity CPU).

The simulation core is a deterministic discrete-event simulator (DES) with a
priority event queue ordered by \emph{(timestamp, event type, sequence ID)}.
This ordering delivers two properties operators need from a fitness function:
\textbf{reproducibility}---identical seeds produce byte-identical outputs, enabling
repeatable algorithm search---and \textbf{conservation}---every injected request
is accounted for as completed, queued, running, or dropped, eliminating simulator
artifact from throughput measurements.

% --------------------------------------------------------------------------
\subsection{Architecture Overview}
\label{subsec:design-arch}

Figure~\ref{fig:blis-arch} shows the full BLIS architecture.
A \code{WorkloadSpec} drives a generator that emits a time-ordered request stream;
requests enter the cluster simulator, which applies admission control, routes each
request to one of $N$ instance simulators, and collects per-request metrics at
completion.
Each instance simulator runs its own DES loop: at every \emph{step} it orders its
wait queue via a pluggable \code{InstanceScheduler}, forms a batch via
\code{BatchFormation}, allocates KV blocks via \code{KVStore}, and advances the
clock by the step duration from \code{LatencyModel}.
The cluster layer introduces \textsc{Admission} and \textsc{Routing} event types
that fire before any instance sees a request, so cluster-level policy decisions
can be recorded, replayed, and swapped independently of per-instance execution.

% ---- Figure A: BLIS Architecture ----
\begin{figure*}[t]
\centering
\begin{tikzpicture}[
  comp/.style={draw, rounded corners=3pt, fill=white,
               minimum width=2.05cm, minimum height=0.50cm,
               align=center, font=\footnotesize, inner sep=2pt},
  plug/.style={draw=gray!55, rounded corners=3pt, fill=white, dashed,
               minimum width=2.05cm, minimum height=0.50cm,
               align=center, font=\footnotesize, inner sep=2pt},
  hdr/.style={font=\small\bfseries, align=center},
  arr/.style={->, >=Stealth, semithick},
  darr/.style={->, >=Stealth, semithick, dashed, green!50!black},
]
%% Background colour bands
\fill[blue!5,   rounded corners=4pt] (-1.35,-2.55) rectangle (1.55, 0.95);
\fill[orange!5, rounded corners=4pt] (1.70,-2.55)  rectangle (10.05,0.95);
\fill[green!5,  rounded corners=4pt] (10.20,-2.55) rectangle (13.45,0.95);

%% Column headers
\node[hdr, blue!65!black]   at (0.10, 1.25) {Workload\\Model};
\node[hdr, orange!70!black] at (5.87, 1.25) {Control-Plane Policies \emph{(pluggable)}};
\node[hdr, green!55!black]  at (11.82,1.25) {Execution\\Physics};

%% ---- Workload column (x≈0.10) ----
\node[comp, fill=blue!12]     (ws)  at (0.10, 0.52) {\textbf{WorkloadSpec}};
\node[comp, fill=blue!5]            at (0.10,-0.08) {arrival process};
\node[comp, fill=blue!5]            at (0.10,-0.60) {token distribution};
\node[comp, fill=blue!5]            at (0.10,-1.12) {SLO class};
\node[comp, fill=blue!5]            at (0.10,-1.64) {cohort dynamics};

%% ---- Admission column (x≈3.10) ----
\node[comp, fill=orange!14]   (adm) at (3.10, 0.52) {Admission};
\node[plug]                         at (3.10,-0.08) {AlwaysAdmit};
\node[plug]                         at (3.10,-0.60) {TokenBucket};

%% ---- Router column (x≈5.87) ----
\node[comp, fill=orange!14]   (rtr) at (5.87, 0.52) {Router};
\node[plug, text width=3.0cm, align=center, font=\tiny, inner sep=3pt]
     at (5.87,-0.42) {prefix-affinity \\ queue-depth \\ kv-utilization \\ load-balance};
\node[font=\tiny, gray]             at (5.87,-1.22) {weighted composite scorer};

%% ---- Instance column (x≈8.64) ----
\node[comp, fill=orange!14, minimum width=2.40cm] (inst) at (8.64, 0.52)
      {Instance ${\times}\,N$};
\node[comp, fill=orange!5]    at (8.64,-0.08) {Scheduler};
\node[comp, fill=orange!5]    at (8.64,-0.60) {Batch formation};
\node[comp, fill=orange!5]    at (8.64,-1.12) {KV Store};
\node[plug, fill=green!8]    (lm) at (8.64,-1.64) {Latency Model $\leftarrow$};

%% ---- Physics column (x≈11.82) ----
\node[plug]                        at (11.82, 0.52) {Blackbox};
\node[plug]                        at (11.82,-0.08) {Roofline};
\node[plug]                        at (11.82,-0.60) {Cross-model};
\node[plug, thick, fill=green!18] (tr) at (11.82,-1.12) {\textbf{Trained-roofline}};
\node[font=\tiny, gray, align=center] at (11.82,-1.72)
      {$\leftarrow$ no data\hspace{1.6em}data-efficient $\rightarrow$};

%% ---- Metrics bar ----
\node[comp, fill=gray!12, minimum width=5.0cm] (met)
      at (5.87,-2.18) {Metrics:\enspace TTFT\enspace TPOT\enspace E2E\enspace SLO};

%% ---- Arrows ----
\draw[arr] (ws.east)   -- node[above,font=\tiny]{requests} (adm.west);
\draw[arr] (adm.east)  -- (rtr.west);
\draw[arr] (rtr.east)  -- (inst.west);
\draw[darr] (tr.west)  -- ++(-0.45,0) |- (lm.east);
\draw[arr]  (inst)     -- ++(0,-0.38) -| (met);
\end{tikzpicture}
\caption{\textbf{BLIS architecture.}
  Three swappable axes: workload model, control-plane policies (dashed borders =
  pluggable implementations), and execution physics.
  Requests flow left to right through admission, routing, and instance simulation to
  metrics.
  The latency model (green) is selected from four backends spanning the
  data-availability spectrum; the trained-roofline backend (\S\ref{subsubsec:trained-roofline})
  achieves 7\% MAPE with 13 production traces.}
\label{fig:blis-arch}
\end{figure*}

% --------------------------------------------------------------------------
\subsection{Learned Execution Physics}
\label{subsec:design-latency}

Step-time error dominates end-to-end prediction error.
A simulator that misestimates batch step durations propagates that error into
every downstream quantity it computes: queue wait times accumulate over step
boundaries, so a 10\% step-time overestimate inflates TTFT, TPOT, and tail-latency
predictions by the same factor.
BLIS offers four latency-model backends, from a zero-data analytical bound to a
data-efficient learned model, all behind a uniform four-method interface
(step time, queueing time, post-decode overhead, per-token processing time).

% ---- Figure B: Roofline to Trained-Roofline ----
\begin{figure}[t]
\centering
\begin{tikzpicture}[scale=0.92]
%% ---- (a) Roofline chart ----
\begin{scope}
  \draw[->] (0,0) -- (4.6,0)
            node[right, font=\scriptsize] {Arith.\ intensity (FLOP/B)};
  \draw[->] (0,0) -- (0,3.4)
            node[above, font=\scriptsize] {Perf.\ (FLOP/s)};

  % Memory-bandwidth slope
  \draw[thick, blue!60!black] (0,0) -- (2.2, 2.75);
  \node[font=\tiny, blue!60!black, rotate=51] at (0.75,1.05) {peak BW};

  % Peak-compute ceiling
  \draw[thick, blue!60!black] (2.2, 2.75) -- (4.4, 2.75);
  \node[font=\tiny, blue!60!black] at (3.5, 2.95) {peak FLOP/s};

  % FP16 ridge
  \draw[densely dashed, gray!70] (2.2,0) -- (2.2,2.75);
  \node[font=\tiny, gray!80, below] at (2.2,-0.05) {\tiny FP16};

  % INT4 ridge (shifted left — quantized weight BW rises)
  \draw[densely dashed, orange!70!black] (1.2,0) -- (1.2,1.50);
  \node[font=\tiny, orange!70!black, below] at (1.2,-0.05) {\tiny INT4};

  % Decode operating point (memory-bound)
  \fill[red!65!black] (0.65, 0.81) circle (2.5pt);
  \node[font=\tiny, above right] at (0.65, 0.81) {decode};

  % Prefill operating point (compute-bound)
  \fill[green!55!black] (3.40, 2.75) circle (2.5pt);
  \node[font=\tiny, below right] at (3.40, 2.75) {prefill};

  \node[font=\small\bfseries] at (2.2, 3.25) {(a) Analytical roofline};
\end{scope}

%% ---- (b) Trained-roofline term decomposition ----
\begin{scope}[xshift=5.4cm]
  \node[font=\small\bfseries] at (1.9, 3.25) {(b) Trained-roofline: term breakdown};

  % Axis
  \draw[->] (0,0)   -- (0,3.0) node[above, font=\scriptsize] {};
  \draw[->] (0,0)   -- (3.9,0) node[right, font=\scriptsize] {};

  % Three batch columns: prefill-heavy, mixed, decode-heavy
  \foreach \x/\lbl in {0.5/prefill-heavy, 1.9/mixed, 3.3/decode-heavy}{
    \node[font=\tiny, below] at (\x, -0.05) {\lbl};
  }

  % Stacked bars: colours for each term
  % Prefill-heavy: T_pf_comp dominates
  \fill[blue!55]    (0.2, 0.00) rectangle (0.8, 1.70); % pf compute
  \fill[green!45]   (0.2, 1.70) rectangle (0.8, 2.00); % weight
  \fill[gray!40]    (0.2, 2.00) rectangle (0.8, 2.25); % fixed

  % Mixed
  \fill[blue!55]    (1.6, 0.00) rectangle (2.2, 0.90); % pf compute
  \fill[red!45]     (1.6, 0.90) rectangle (2.2, 1.65); % dc kv
  \fill[green!45]   (1.6, 1.65) rectangle (2.2, 2.00); % weight
  \fill[gray!40]    (1.6, 2.00) rectangle (2.2, 2.25); % fixed

  % Decode-heavy: T_dc_kv dominates
  \fill[red!45]     (3.0, 0.00) rectangle (3.6, 1.80); % dc kv
  \fill[green!45]   (3.0, 1.80) rectangle (3.6, 2.10); % weight
  \fill[gray!40]    (3.0, 2.10) rectangle (3.6, 2.35); % fixed

  % Legend
  \fill[blue!55]  (0.0, 2.65) rectangle (0.25, 2.85);
  \node[font=\tiny, right] at (0.28, 2.75) {$\beta_1 T^\text{pf}$ (prefill)};
  \fill[red!45]   (0.0, 2.40) rectangle (0.25, 2.60);
  \node[font=\tiny, right] at (0.28, 2.50) {$\beta_2 T^\text{dc}$ (decode)};
  \fill[green!45] (2.0, 2.65) rectangle (2.25,2.85);
  \node[font=\tiny, right] at (2.28, 2.75) {$\beta_3 T_\text{weight}$};
  \fill[gray!40]  (2.0, 2.40) rectangle (2.25,2.60);
  \node[font=\tiny, right] at (2.28, 2.50) {$\beta_5 L + \beta_6 B + \beta_7$};
\end{scope}
\end{tikzpicture}
\caption{\textbf{Latency-model backends.}
  (a)~The analytical roofline places prefill in the compute-bound regime
  and decode in the memory-bandwidth-bound regime.
  Quantization shifts the roofline ridge left (INT4 vs.\ FP16), reducing KV-cache
  pressure and moving the decode point closer to the ridge.
  (b)~The trained-roofline decomposes step time into four physically-motivated terms
  whose relative contributions shift with batch composition.}
\label{fig:latency-model}
\end{figure}

\subsubsection{The Analytical Roofline Model}
\label{subsubsec:roofline}

The roofline model bounds step time by the slower of compute throughput and memory
bandwidth:
\begin{equation}
  T_\text{step} = \max\!\left(\frac{F}{\hat{P}_\text{flops}},\;\frac{M}{\hat{P}_\text{bw}}\right),
  \label{eq:roofline}
\end{equation}
where $\hat{P}_\text{flops} = P_\text{flops} \times \mathrm{MFU}$ and
$\hat{P}_\text{bw} = P_\text{bw}$ are the hardware's effective FLOP/s and byte/s
throughput.
BLIS computes $F$ from first principles for each batch: QKV projection, attention
score computation, output projection, and MLP (two-matrix SwiGLU convention),
with each term differentiated by whether tokens are in the prefill or decode phase.
The memory term $M$ covers three sources: model weight bytes loaded once per step,
KV-cache blocks written during prefill, and KV-cache blocks read and written during
decode.

The model uses \emph{phase-specific} MFU correction factors (\code{MfuPrefill},
\code{MfuDecode}) because prefill and decode operate at fundamentally different
arithmetic intensities (Figure~\ref{fig:latency-model}a).
For quantized models, BLIS uses \code{EffectiveWeightBytesPerParam()}---auto-detected
from HuggingFace quantization config, model-name conventions (e.g., \code{w4a16}
$\rightarrow$ 0.5 bytes/param), or the \code{torch\_dtype} fallback---for the weight
term while keeping the full-precision byte count for KV-cache and activation memory.
This correctly shifts the roofline ridge for quantized models without requiring
separate profiling.

\subsubsection{The Trained-Roofline Model}
\label{subsubsec:trained-roofline}

The analytical roofline captures the right \emph{structure} but the wrong
\emph{scale}: hardware non-idealities---NCCL synchronization, scheduler overhead,
memory system effects---shift the contribution of each roofline term by a
model-specific factor that no analytical formula predicts.
The trained-roofline model promotes each term to a learnable basis function with a
fitted scalar correction:
\begin{equation}
\boxed{
  T_\text{step} =
    \beta_1\max(T^\text{pf}_\text{comp},\,T^\text{pf}_\text{kv})
  + \beta_2\max(T^\text{dc}_\text{comp},\,T^\text{dc}_\text{kv})
  + \beta_3\,T_\text{weight}
  + \beta_4\,T_\text{tp}
  + \beta_5\,L
  + \beta_6\,B
  + \beta_7.
}
  \label{eq:trained-roofline}
\end{equation}
Each term has a precise physical interpretation.
$T^\text{pf}_\text{comp}$ and $T^\text{pf}_\text{kv}$ are the prefill
compute time and KV-write bandwidth time; their max captures the phase-level
bottleneck.
$T^\text{dc}_\text{comp}$ and $T^\text{dc}_\text{kv}$ are the analogous
decode terms.
$T_\text{weight}$ is the weight-loading time, amortized over the batch.
$T_\text{tp} \approx 0$ on H100 NVLink (TP communication is absorbed into the
per-layer overhead $\beta_5 L$).
$B$ is the batch size.

BLIS fits these seven coefficients from 13 vLLM experiments spanning four
architectures (Llama-3, Qwen3, DeepSeek-MoE, Granite):
\[
  \bm{\beta} = [0.773,\;1.127,\;1.056,\;0.0,\;43.5\,\mu\text{s},\;48.8\,\mu\text{s},\;0.0].
\]
That $\beta_1$--$\beta_3$ are near 1.0 is not coincidental: the roofline basis
captures the right physics, and the corrections are small rather than compensatory.
Figure~\ref{fig:latency-model}b shows how term contributions shift across
prefill-heavy, mixed, and decode-heavy batches.
For MoE models, the weight term uses $n_\text{eff} = \min(N,\max(k,\,B{\cdot}k))$
expert loads per step, modelling the routing fan-out without additional coefficients.

\subsubsection{End-to-End Request Latency Decomposition}
\label{subsubsec:e2e-decomp}

End-to-end latency decomposes as
\begin{equation}
  \mathrm{E2E}(r) \;=\;
  \underbrace{\alpha_0}_{\text{API overhead}}
  + \sum_{s \,\in\, \mathrm{steps}(r)} T_s
  + \underbrace{\alpha_1}_{\text{post-decode}}
  + \underbrace{\alpha_2 \cdot y_r}_{\text{detokenization}},
  \label{eq:e2e}
\end{equation}
where $y_r$ is the number of output tokens.
TTFT is the time from arrival to the first decode step; TPOT is the mean
inter-token interval during decode.
For the trained-roofline backend, $\alpha_0$ is a constant API overhead (the
prefill cost is already captured in step-time terms), $\alpha_1$ is a fixed
post-decode cost, and $\alpha_2$ is the per-token detokenization latency.
For the blackbox and cross-model backends, $\alpha_0$ grows linearly with
input length, modelling the end-to-end queueing and KV-prefill overhead observed
in trace data.

\subsubsection{Training Pipeline}
\label{subsubsec:training}

\begin{sloppypar}
BLIS fits the trained-roofline coefficients in three sequential phases using
non-negative least squares (NNLS).
Phase~1 isolates $\beta_3$ from batches where $T_\text{weight}$ dominates
(zero prefill and decode tokens).
Phase~2 fixes $\beta_3$ and fits $\beta_1$, $\beta_2$ from mixed batches that
vary prefill-to-decode ratio.
Phase~3 fits $\beta_5$, $\beta_6$, $\beta_7$ from the step-time residual.
This three-phase structure exploits physical separability: earlier
coefficients are not disturbed by later phases.
Across 137{,}000 requests in 13 experiments, the trained-roofline achieves
\textbf{7\% MAPE} (mean absolute percentage error) in step-time prediction.
The physical grounding is also what makes the model data-efficient: 13 experiments
across four architectures suffice because the basis functions are right.
A purely data-driven polynomial fit requires far more data to generalize across
held-out architectures, since its basis functions must be discovered rather than
derived from physics.
\end{sloppypar}

% --------------------------------------------------------------------------
\subsection{Pluggable Control-Plane Policies}
\label{subsec:design-policies}

The control plane of an LLM serving system is deterministic given its inputs:
routing decisions, scheduling orders, and admission outcomes depend only on
observable queue state, not on GPU execution.
BLIS replicates each control-plane mechanism behind a single-method interface,
making every policy independently swappable and testable.
Faithful replication matters: a routing policy that silently ignores prefix
affinity causes cache misses that inflate step times, conflating two distinct
sources of error that must be diagnosed and corrected independently.

\subsubsection{Admission Control}
\label{subsubsec:admission}

Admission fires at the cluster level before any instance sees a request.
The \code{AdmissionPolicy} interface takes a request and the current cluster
state (\code{RouterState}) and returns an admit/reject decision with a reason.
BLIS ships two implementations: \textsc{AlwaysAdmit} (the default,
suitable for overload-free regimes) and \textsc{TokenBucket} (rate-limiting
with configurable capacity and refill rate, cost proportional to input-token
count).
Because rejection happens before routing, dropped requests never consume
instance queue state, enabling clean overload experiments where the admission
threshold is the variable under study.

\subsubsection{Request Routing}
\label{subsubsec:routing}

\begin{sloppypar}
BLIS supports four routing policies: \textsc{RoundRobin},
\textsc{LeastLoaded}, \textsc{WeightedScoring}, and a pathological
\textsc{AlwaysBusiest} baseline.
The real contribution is the \emph{composable scorer framework} inside
\textsc{WeightedScoring} (Figure~\ref{fig:scorer}).
Pure load-balancing ignores prefix locality and incurs cold-cache misses on
every step; pure affinity routing ignores load imbalance and overloads
hot-prefix instances.
The composable scorer resolves this tension: each scorer encodes one signal,
and operators control the tradeoff with a weight vector.
Each scorer maps every candidate instance to a score in $[0,1]$.
The composite score is $\sum_i w_i \cdot s_i$ with normalised weights;
the instance with the highest composite score is selected, with random
tie-breaking for reproducibility.
\end{sloppypar}

BLIS provides four built-in scorers.
\emph{prefix-affinity} scores by prefix-match ratio, using a router-side LRU
index that tracks block hashes per instance---the router maintains a shadow of
each instance's KV-cache topology without reading engine internals.
\emph{queue-depth} uses min-max normalised effective load.
\emph{kv-utilization} scores as $1 - \mathrm{KVUtil}$.
\emph{load-balance} uses an inverse-load transform that preserves absolute
load differences.
The default profile \code{prefix-affinity:3,\,queue-depth:2,\,kv-utilization:2}
matches the llm-d endpoint picker~\cite{llmd2024}, enabling direct sim-to-real
comparison without re-tuning weights.

% ---- Figure C: Composable Scorer Framework ----
\begin{figure}[t]
\centering
\begin{tikzpicture}[
  inst/.style={draw, rounded corners=3pt, fill=gray!8,
               minimum width=1.55cm, minimum height=0.45cm,
               align=center, font=\footnotesize},
  arr/.style={->, >=Stealth, semithick},
  lbl/.style={font=\tiny},
]
% Three instances across the top
\foreach \i/\x in {1/0.0, 2/2.2, 3/4.4}{
  \node[inst] (I\i) at (\x, 2.8) {Instance$_\i$};
}

% Stacked score bars for each instance
% Colours: blue=prefix-affinity, orange=queue-depth, teal=kv-util
\foreach \x/\pa/\qd/\kv in {0.0/1.40/0.70/0.45, 2.2/0.30/1.05/0.85, 4.4/0.85/0.60/0.60}{
  \pgfmathsetmacro{\base}{0.0}
  \fill[blue!50]   (\x-0.55, \base)           rectangle (\x+0.55, \base+\pa);
  \fill[orange!60] (\x-0.55, \base+\pa)       rectangle (\x+0.55, \base+\pa+\qd);
  \fill[teal!50]   (\x-0.55, \base+\pa+\qd)   rectangle (\x+0.55, \base+\pa+\qd+\kv);
  \draw            (\x-0.55, 0) rectangle (\x+0.55, \base+\pa+\qd+\kv);
}

% Score axis label
\draw[->] (-0.85,0) -- (-0.85, 2.75) node[above, font=\tiny] {score};
\foreach \y in {0,0.5,1.0,1.5,2.0,2.5}{
  \draw[gray!40] (-0.90,\y) -- (5.0,\y);
}

% Argmax marker
\draw[arr, red!70!black, very thick] (0.0, 2.60) -- (0.0, 2.82);
\node[font=\tiny, red!70!black, above] at (0.0, 2.82) {\textbf{selected}};

% Legend — weights shown next to each scorer name
\fill[blue!50]   (5.4, 2.4) rectangle (5.7, 2.6);
\node[font=\tiny, right] at (5.72, 2.50) {prefix-affinity\ ($w{=}3$)};
\fill[orange!60] (5.4, 2.1) rectangle (5.7, 2.3);
\node[font=\tiny, right] at (5.72, 2.20) {queue-depth\ ($w{=}2$)};
\fill[teal!50]   (5.4, 1.8) rectangle (5.7, 2.0);
\node[font=\tiny, right] at (5.72, 1.90) {kv-utilization\ ($w{=}2$)};

% Composite formula note
\node[font=\tiny, gray, align=center] at (2.2, -0.45)
      {composite $= \textstyle\sum_i w_i \cdot s_i$,\enspace $s_i \in [0,1]$};
\end{tikzpicture}
\caption{\textbf{Composable scorer framework.}
  Three scorers evaluate all candidate instances independently on $[0,1]$.
  Weighted combination selects Instance$_1$ here because its high prefix-affinity
  score (large blue bar) outweighs lower load-balance scores.
  The default weights (pa:3, qd:2, kv:2) match the llm-d endpoint picker,
  enabling direct sim-to-real comparison.
  New scorers plug in by implementing a single-method interface.}
\label{fig:scorer}
\end{figure}

\subsubsection{Instance Scheduling and Batch Formation}
\label{subsubsec:scheduling}

Inside each instance, \code{InstanceScheduler} re-orders the wait queue before
every step.
BLIS provides three meaningful scheduling policies: \textsc{FCFS} (default),
\textsc{PriorityFCFS} (sort by SLO class then arrival time), and
\textsc{SJF} (sort by input-token count).
Priority values come from a separate \code{PriorityPolicy}; the
\textsc{SLOBased} variant uses $\text{base} + w_\text{age} \times \text{wait}$,
where the aging term prevents starvation across SLO classes.

Batch formation replicates vLLM's continuous batching with chunked prefill and
preemption.
Phase~1 processes continuing requests: long prompts are chunked to
\code{LongPrefillTokenThreshold} tokens per step, and decode tokens fill the
remainder of the \code{MaxScheduledTokens} budget.
Phase~2 admits new requests from the wait queue, allocating KV blocks and
looking up prefix-cached blocks.
If KV allocation fails during Phase~1, the formation preempts from the batch
tail---preserving earlier blocks that more requests may share---until memory
is available or the batch reaches size one.
This two-phase algorithm is the same structure used in production vLLM, ensuring
that BLIS batch dynamics match real deployments.

\subsubsection{KV Cache and Memory Tiering}
\label{subsubsec:kvcache}

BLIS models KV-cache pressure at block granularity.
The single-tier GPU cache maintains an O(1) LRU free list and a hash-to-block
index for prefix reuse.
Block hashes use a hierarchical SHA-256 chain: each block's hash includes the
previous block's hash, so the prefix-match lookup costs $O(K \cdot B_\text{sz})$
rather than $O(K^2 \cdot B_\text{sz})$, where $K$ is the cached-prefix length
in blocks and $B_\text{sz}$ is the block size.
A failed allocation rolls back atomically---restoring the free list, hash index,
and block states---so the batch formation can retry or preempt without leaving
the cache in an inconsistent state.

The tiered GPU\,+\,CPU backend (\code{TieredKVCache}) composes the GPU tier with
a CPU LRU mirror (Figure~\ref{fig:tiered-kv}).
After each batch step, \code{MirrorToCPU} copies every completed full block's
hash and token data to the CPU---lazy mirroring rather than eviction-triggered
transfer.
When GPU allocation fails on a cache miss, \code{reloadPrefixFromCPU} walks the
prefix hash chain, pops a GPU free block for each CPU hit, and accumulates a PCIe
transfer stall of $t_\text{base} + \lceil B_\text{sz} / \text{BW}_\text{PCIe} \rceil$
per reloaded block.
The stall is injected into the next step via \code{ConsumePendingTransferLatency},
modelling the real system's PCIe serialisation without cycle-accurate simulation.
KV capacity is computed analytically: total GPU memory minus model weights
(three-matrix SwiGLU, consistent with the trained-roofline convention),
activation memory, and non-torch overhead, divided by bytes-per-block at the
configured precision.

% ---- Figure D: Tiered KV Cache ----
\begin{figure}[t]
\centering
\begin{tikzpicture}[
  mem/.style={draw, rounded corners=3pt, minimum height=0.80cm,
              minimum width=5.6cm, align=center, font=\small},
  blk/.style={draw, minimum width=0.62cm, minimum height=0.62cm,
              font=\tiny, align=center},
  arr/.style={->, >=Stealth, semithick},
]
%% GPU tier — raised to y=2.7 to open up the inter-tier gap
\node[mem, fill=blue!8]  (gpu) at (0, 2.7) {};
\node[font=\small\bfseries, blue!60!black, left=0.1cm] at (gpu.west) {GPU VRAM};

\foreach \i/\col/\lab in {
  1/blue!25/{A}, 2/blue!25/{B}, 3/orange!30/{C},
  4/green!25/{D}, 5/gray!20/{free}, 6/gray!20/{free}}{
  \node[blk, fill=\col] at (-2.1 + \i*0.75, 2.7) {\lab};
}

%% CPU tier at y=0.5
\node[mem, fill=orange!6] (cpu) at (0, 0.5) {};
\node[font=\small\bfseries, orange!70!black, left=0.1cm] at (cpu.west) {CPU DRAM};

\foreach \i/\col/\lab in {
  1/blue!15/{A}, 2/blue!15/{B}, 3/green!15/{D},
  4/gray!15/{E}, 5/gray!15/{F}, 6/gray!15/{G}}{
  \node[blk, fill=\col] at (-2.1 + \i*0.75, 0.5) {\lab};
}

%% Inter-tier gap: GPU bottom = 2.30, CPU top = 0.90
%% Arrows to the right of all named blocks (x > 2.71)
\draw[arr, blue!55!black]   (3.4, 2.30) -- (3.4, 0.90)
    node[midway, right, font=\tiny, blue!55!black] {lazy mirror};

\draw[arr, dashed, orange!65!black] (3.0, 0.90) -- (3.0, 2.30)
    node[midway, right, font=\tiny, orange!65!black] {reload on miss};

%% PCIe stall — below CPU box (CPU box bottom = 0.10)
\node[draw=orange!50, rounded corners=2pt, fill=orange!5,
      font=\tiny, align=center, inner sep=2pt]
     at (3.2, -0.28) {PCIe: $t_\text{base}+\lceil B_\text{sz}/\mathrm{BW}\rceil$\\per block};
\draw[->, very thin, dashed, orange!45!black] (3.2, -0.08) -- (3.2, 0.87);
\end{tikzpicture}
\caption{\textbf{Tiered KV cache.}
  After each step, BLIS lazily mirrors newly completed blocks from GPU to CPU DRAM.
  On a GPU miss, matching CPU blocks reload to GPU and accumulate a PCIe transfer
  stall injected into the next step's execution time.
  Orange block~C is on GPU but not yet mirrored; green block~D has been mirrored to both tiers.}
\label{fig:tiered-kv}
\end{figure}

% --------------------------------------------------------------------------
\subsection{Composable Workload Models}
\label{subsec:design-workloads}

Workload fidelity matters as much as execution-physics accuracy: a simulator fed
a naive Poisson stream with fixed token lengths will predict a deployment that
no real workload ever produces.
ServeGen~\cite{xiang2025servegen} shows that per-client compositional generation
avoids the 50\% under-provisioning errors of aggregate workload generation;
BLIS's workload model follows the same per-client structure.

A \code{WorkloadSpec} organises parameters at three levels.
At the \emph{cluster level}: aggregate request rate, simulation horizon, and RNG seed.
At the \emph{client level}: each client specifies an arrival process (four
families, from Poisson to heavy-tailed Weibull), token-length distributions
(five families, including an empirical type that replays a ServeGen trace with
zero re-parameterisation~\cite{xiang2025servegen}), SLO class, prefix group,
and session mode (open-loop or closed-loop with configurable think time).
At the \emph{cohort level}: population dynamics---diurnal scaling, traffic spikes,
client drain---produce the bursty, non-stationary arrival patterns observed in
production.
For closed-loop sessions the generator emits only round-0 requests at
initialisation; subsequent rounds are generated at runtime as completions arrive,
preserving the causal dependency between turns.

% --------------------------------------------------------------------------
\subsection{Metrics and Fitness Evaluation}
\label{subsec:design-metrics}

BLIS measures three per-request latency quantities operators care about:
\emph{time to first token} (TTFT), \emph{time per output token} (TPOT), and
\emph{end-to-end latency} (E2E, Equation~\ref{eq:e2e}).
For each, it reports the full empirical distribution and its percentiles (p50,
p95, p99), along with per-SLO-class breakdowns, aggregate throughput (tokens/s
and requests/s), and KV-cache hit rate.

For algorithm evolution (\S\ref{sec:eval-evolution}), BLIS exposes a scalar
fitness function:
\begin{equation}
  \mathcal{F}(\pi,\,\mathcal{W}) \;=\; -\,\mathrm{p95}(\mathrm{E2E})
  \quad\text{s.t.}\quad
  \mathrm{p99}(\mathrm{TTFT}) \;\leq\; T_\mathrm{slo},
  \label{eq:fitness}
\end{equation}
The objective minimises p95(E2E)---the experience of the slowest one-in-twenty
users---while the constraint prevents the optimizer from sacrificing first-token
responsiveness for throughput.
$\pi$ is a policy configuration (routing weights, scheduling policy,
admission threshold) and $\mathcal{W}$ is a workload scenario.
A single simulation evaluates $\mathcal{F}$ in under two seconds on a laptop CPU,
making 1{,}000-candidate evolutionary search feasible in under 30 minutes---orders
of magnitude faster than real-hardware evaluation, which requires minutes of GPU
time per configuration.

% ============================================================
% SECTION 5: SIM-TO-REAL METHODOLOGY
% ============================================================
\section{Sim-to-Real Transfer Methodology}
\label{sec:simtoreal}

(To be written: experimental protocol for sim-to-real validation.)

\subsection{Real-System Testbed}
\label{subsec:simtoreal-testbed}
\subsection{Calibration Protocol}
\label{subsec:simtoreal-calibration}
\subsection{Transfer Validation Protocol}
\label{subsec:simtoreal-transfer}

% ============================================================
% SECTION 6: EVALUATION I --- PREDICTION ACCURACY
% ============================================================
\section{Prediction Accuracy}
\label{sec:eval-accuracy}

(To be written: BLIS vs.\ Vidur, InferSim, BentoML, TensorRT-LLM.)

\subsection{Baselines}
\label{subsec:accuracy-baselines}
\subsection{Step-Time Prediction}
\label{subsec:accuracy-steptime}
\subsection{End-to-End Latency Under Fixed Configuration}
\label{subsec:accuracy-e2e-fixed}
\subsection{Latency Under Configuration Variation}
\label{subsec:accuracy-e2e-varied}
\subsection{Accuracy Under Workload Shift}
\label{subsec:accuracy-workload}

% ============================================================
% SECTION 7: EVALUATION II --- CAPACITY PLANNING AND CONFIG SEARCH
% ============================================================
\section{Capacity Planning and Configuration Search}
\label{sec:eval-capplan}

(To be written: case studies with sim-to-real validation.)

\subsection{Problem Formulation}
\label{subsec:capplan-formulation}
\subsection{Case Study 1: GPU Count Provisioning}
\label{subsec:capplan-gpucount}
\subsection{Case Study 2: Serving Configuration Optimization}
\label{subsec:capplan-config}
\subsection{Case Study 3: KV Cache Tier Sizing}
\label{subsec:capplan-kvsizing}
\subsection{Decision Quality Comparison}
\label{subsec:capplan-decision}

% ============================================================
% SECTION 8: EVALUATION III --- ALGORITHM EVOLUTION
% ============================================================
\section{Algorithm Evolution via AI-Native Systems}
\label{sec:eval-evolution}

(To be written: evolve routing/scheduling policies via ADRS, transfer to production.)

\subsection{ADRS Methodology and BLIS as Fitness Substrate}
\label{subsec:evolution-adrs}
\subsection{Evolving Routing Policies}
\label{subsec:evolution-routing}
\subsection{Evolving Scheduling Strategies}
\label{subsec:evolution-scheduling}
\subsection{Evolving Admission Control Under Load Spikes}
\label{subsec:evolution-admission}
\subsection{Sim-to-Real Transfer of Evolved Algorithms}
\label{subsec:evolution-transfer}

% ============================================================
% SECTION 9: DISCUSSION
% ============================================================
\section{Discussion}
\label{sec:discussion}

(To be written.)

\subsection{Limitations and When BLIS Falls Short}
\label{subsec:discuss-limitations}
\subsection{Generalization to Other Inference Engines}
\label{subsec:discuss-generalization}
\subsection{Toward Continuous Calibration}
\label{subsec:discuss-continuous}
\subsection{BLIS as Infrastructure for Systems Research}
\label{subsec:discuss-research}

% ============================================================
% SECTION 10: CONCLUSION
% ============================================================
\section{Conclusion}
\label{sec:conclusion}

(To be written.)


% ============================================================
% APPENDICES
% ============================================================
\begin{appendices}

% -------------------------------------------------------
\section{Step-Time Basis Function Derivations}
\label{app:basis}

% TODO: Full mathematical derivations for all 7 basis functions
(To be written.)

% -------------------------------------------------------
\section{Training Pipeline Details}
\label{app:training}

% TODO: Extended details on the training pipeline
(To be written.)

% -------------------------------------------------------
\section{Dataset Description}
\label{app:dataset}

% TODO: Full description of the 16 experiments (13 active + 3 overload)
(To be written.)

% -------------------------------------------------------
\section{Extended Evaluation Results}
\label{app:extended-eval}

% TODO: Additional experimental results
(To be written.)

% -------------------------------------------------------
\section{Reproducibility}
\label{app:reproducibility}

% TODO: Detailed instructions for reproducing all experiments
(To be written.)

\end{appendices}


\bibliographystyle{plain}
\bibliography{blis}

\end{document}
